---
title: "Final_Project_Binpeng_Zhang"
author: "Binpeng Zhang"
date: "5/14/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Pokemon Tutorial: How To Be The Very Best Dragon Pokemon Trainer
# Introduction:
Since Pokemon's release in 1996, it has grown in popularity. There have been many animes, games and trading cards based on the series. The immense growth in popular and a huge fanbase allowed for Pokemon to become a competitive field. There are many people dedicated to becoming the very best pokemon trainer in the world. I have created this simple and easier to understand tutorial that will teach you to become the very best dragon type pokemon trainer, because dragons are the best. I will also teach you Data curation, parsing and management, exploratory data analysis, hypothesis testing, machine learning and how to draw insights from the data set using R. 
## Data Curation, Parsing and Management
First we have to load in the dataset of 721 Pokemon, including their number, name, first and second type, and basic stats: HP, Attack, Defense, Special Attack, Special Defense, and Speed. We will also load the necessary libraries such as tidyverse and stringr that will help us later with function calls. 

```{r Data Curation}
library(tidyverse)
library(stringr)
PokemonStats <- read.csv("/Users/bzhang/Documents/cmsc320/Project4/Pokemon.csv")
head(PokemonStats)
```
Since we only want to be a Dragon Pokemon Trainer we will tidy the data to filter out all useless non dragon type pokemons. We will also tidy names of pokemon such as AltariaMega Altaria to just AltariaMega since repeating Altaria is just redundant. 
```{r Tidy Data}
DragonStats <- PokemonStats %>% 
  filter(Type.1 == "Dragon")
# before tidying Mega
DragonStats
DragonStats$Name<- str_extract(DragonStats$Name,boundary("word"))
# after tidying all the Mega pokemons
DragonStats
```
In some competitive tournaments legendary pokemons are not allowed so I will create two data sets: 
One data set for Legendary Pokemons
One data set for Non-Legendary Pokemons
```{r legendary}
LegendaryDragons <- DragonStats %>%
  filter(Legendary == "True")

NonLegendary <- DragonStats %>%
  filter(Legendary == "False")

LegendaryDragons
NonLegendary
```

## Exploratory Data Analysis
Now that we have tidied the data to what we are interested in we can perform some exploratory data analysis on it. We can do things like finding the strongest legendary dragon pokemon or the strongest non legendary pokemon. We define the strongest by the pokemon with the highest total stat. We can find the strongest legendary dragon pokemon by arranging the LegendaryDragons data frame in descending order by total stats using the arrange function and taking the first entity. 
```{r strongestlegendary}
StrongestLegend <- LegendaryDragons %>%
  arrange(desc(Total)) %>%
  head(1)
StrongestLegend
```

We can find the strongest non legendary dragon pokemon by arranging the NonLegendary data frame in descending order by total stats using the arrange function and taking the first entity. 
```{r strongestnonlegend}
StrongestNonLegend <- NonLegendary %>%
  arrange(desc(Total)) %>%
  head(1)
StrongestNonLegend
```
As we can see from the two data frame, the strongest legendary dragon pokemon RayquazaMegais stronger than the strongest non legendary dragon pokemon SalamenceMega by 80 total stat points. This may suggest that legendary dragon pokemons are significantly stronger than non legendary dragon pokemons. 
We will now create some data frames that will summarize these two dataset that will help us get a better analysis of what's going on. 
```{r summaries}
NonLegendarySummary <- DragonStats %>% 
  group_by(Legendary) %>%
  summarize(min_stat=min(Total), mean_stat=mean(Total), max_stat=max(Total) , sd_stat = sd(Total))
NonLegendarySummary
```
From the above table it seems that the average legendary dragon pokemon is much stronger than the average non legendary dragon pokemon. Now we want to visualize the data so we can have a pictorial representation of what's going on so it is more intuitive. 

```{r grapphs}
DragonStats %>%
  ggplot(mapping = aes(x=Name,y=Total)) +
  geom_jitter(aes(color = Legendary)) +
  ggtitle("Legendary vs Non-Legendary") + 
  ylab("Total Stats") + xlab("Name")
```
From the scatter plot above it seems pretty obvious that legendary dragon pokemons are stronger in general, however just to be even more intuitive we are going to create a box plot to have a direct comparison.
```{r boxplot}
boxplot(Total~Legendary,data=DragonStats, main="Legendary vs Non-Legendary", 
  	xlab="Legendary?", ylab="Total Stats")
```
By observing this boxplot it is obvious that in general legendary dragon pokemons are stronger than non legendary pokemons almost 100% of the legendary dragon pokemons are stronger than 75% of the non legendary dragon pokemons. 
## Hypothesis Testing and Machine Learning
So from the boxplot it's very apparent that legendary dragon pokemons are stronger, so I want to do a hypothesis test on whether or not being legendary has a significant impact. The null hypothesis is that there is no significant difference between Legendary Dragon Pokemons and Non-Legendary Dragon Pokemons.  
```{r hypothesistesting}
lgm <- lm(Total~Legendary, data=DragonStats)
broom::tidy(lgm)
```
I used a linear regression line to check if being a legendary dragon pokemon has a significant impact on the total stats. Then I used the tidy function from the broom library to create a table. From observing the p.value of the tidied table, I will have to reject the null hypothesis that there is no significant difference between the total stats of Legendary Dragon Pokemons and Non-Legendary Dragon Pokemons. 
## Insights