---
title: "Final_Project_Binpeng_Zhang"
author: "Binpeng Zhang"
date: "5/14/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Pokemon Tutorial: How To Be The Very Best Dragon Pokemon Trainer
# Introduction:
Since Pokemon's release in 1996, it has grown in popularity. There have been many animes, games and trading cards based on the series. The immense growth in popular and a huge fanbase allowed for Pokemon to become a competitive field. There are many people dedicated to becoming the very best pokemon trainer in the world. I have created this simple and easier to understand tutorial that will teach you to become the very best dragon type pokemon trainer, because dragons are the best. I will also teach you Data curation, parsing and management, exploratory data analysis, hypothesis testing, machine learning and how to draw insights from the data set using R. 
## Data Curation, Parsing and Management
First we have to load in the dataset of 721 Pokemon, including their number, name, first and second type, and basic stats: HP, Attack, Defense, Special Attack, Special Defense, and Speed from https://www.kaggle.com/abcsds/pokemon. We will also load the necessary libraries such as tidyverse and stringr that will help us later with function calls. 
You can get more information on the libraries by clicking the following links:
  -https://www.tidyverse.org/ for information on the tidyverse library
  -https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html for information on the stringr library

```{r Data Curation}
library(tidyverse)
library(stringr)
PokemonStats <- read.csv("/Users/bzhang/Documents/cmsc320/Project4/Pokemon.csv")
head(PokemonStats)
```
Since we only want to be a Dragon Pokemon Trainer we will tidy the data to filter out all useless non dragon type pokemons. We will also tidy names of pokemon such as AltariaMega Altaria to just AltariaMega since repeating Altaria is just redundant. 
```{r Tidy Data}
DragonStats <- PokemonStats %>% 
  filter(Type.1 == "Dragon" | Type.2 == "Dragon")
# before tidying Mega
DragonStats
DragonStats$Name<- str_extract(DragonStats$Name,boundary("word"))
# after tidying all the Mega pokemons
DragonStats
# select only the fields I am interested in
DragonStats <- select(DragonStats, X. , Name, Type.1,Type.2,Total, Generation, Legendary)
DragonStats
```
In some competitive tournaments legendary pokemons are not allowed so I will create two data sets: 
One data set for Legendary Pokemons
One data set for Non-Legendary Pokemons
```{r legendary}
LegendaryDragons <- DragonStats %>%
  filter(Legendary == "True")

NonLegendary <- DragonStats %>%
  filter(Legendary == "False")

LegendaryDragons
NonLegendary
```

## Exploratory Data Analysis
Now that we have tidied the data to what we are interested in we can perform some exploratory data analysis on it. We can do things like finding the strongest legendary dragon pokemon or the strongest non legendary pokemon. We define the strongest by the pokemon with the highest total stat. We can find the strongest legendary dragon pokemon by arranging the LegendaryDragons data frame in descending order by total stats using the arrange function and taking the first entity. 
```{r strongestlegendary}
StrongestLegend <- LegendaryDragons %>%
  arrange(desc(Total)) %>%
  head(1)
StrongestLegend
```

We can find the strongest non legendary dragon pokemon by arranging the NonLegendary data frame in descending order by total stats using the arrange function and taking the first entity. 
```{r strongestnonlegend}
StrongestNonLegend <- NonLegendary %>%
  arrange(desc(Total)) %>%
  head(1)
StrongestNonLegend
```
As we can see from the two data frame, the strongest legendary dragon pokemon RayquazaMegais stronger than the strongest non legendary dragon pokemon SalamenceMega by 80 total stat points. This may suggest that legendary dragon pokemons are significantly stronger than non legendary dragon pokemons. 
We will now create some data frames that will summarize these two dataset that will help us get a better analysis of what's going on. 
```{r summaries}
NonLegendarySummary <- DragonStats %>% 
  group_by(Legendary) %>%
  summarize(min_stat=min(Total), mean_stat=mean(Total), max_stat=max(Total) , sd_stat = sd(Total))
NonLegendarySummary
```
From the above table it seems that the average legendary dragon pokemon is much stronger than the average non legendary dragon pokemon. Now we want to visualize the data so we can have a pictorial representation of what's going on so it is more intuitive. 

```{r grapphs}
nonvslegend <- DragonStats %>%
  ggplot(mapping = aes(x=Name,y=Total)) +
  geom_point(aes(color = Legendary)) +
  ggtitle("Legendary vs Non-Legendary") + 
  ylab("Total Stats") + xlab("Name")
nonvslegend
```
From the scatter plot above it seems pretty obvious that legendary dragon pokemons are stronger in general, however just to be even more intuitive we are going to create a box plot to have a direct comparison.
```{r boxplot}
b_plot <- boxplot(Total~Legendary,data=DragonStats, main="Legendary vs Non-Legendary", xlab="Legendary?", ylab="Total Stats")
b_plot
```
By observing this boxplot it is obvious that in general legendary dragon pokemons are stronger than non legendary pokemons almost 100% of the legendary dragon pokemons are stronger than 75% of the non legendary dragon pokemons. 
## Hypothesis Testing and Machine Learning
For more information on how to do hypothesis testing you can visit this link:
http://www.r-tutor.com/elementary-statistics/hypothesis-testing

So from the boxplot it's very apparent that legendary dragon pokemons are stronger, so I want to do a hypothesis test on whether or not being legendary has a significant impact. The null hypothesis is that there is no significant difference between Legendary Dragon Pokemons and Non-Legendary Dragon Pokemons.  
You can get more information on the broom library by going here: https://cran.r-project.org/web/packages/broom/vignettes/broom.html
```{r hypothesistesting}
lgm <- lm(Total~Legendary, data=DragonStats)
broom::tidy(lgm)
```
I used a linear regression line to check if being a legendary dragon pokemon has a significant impact on the total stats. Then I used the tidy function from the broom library to create a table. From observing the p.value of the tidied table, I will have to reject the null hypothesis that there is no significant difference between the total stats of Legendary Dragon Pokemons and Non-Legendary Dragon Pokemons because the p.value is 4.280380e-07 which is less than the significance value of 0.05. 

Okay, now that we can conclude that legendary dragon pokemons are stronger, let's do something a bit harder. For example, maybe you want to be a dragon generation pokemon enthusiast who wants to use pokemons of only one generations. We can accomplish this by grouping by the generations. We are going to make three data sets one for the legendaries, one for the non-legendaries and one for both so we can have a comparison. 
```{r generations}
#legendary generation boxplot
l_gen <- boxplot(Total~Generation,data=LegendaryDragons, main="Legendary Dragon Boxplot for Each Generations",xlab="Generation", ylab="Total Stats")
l_gen

#non-legendary generation boxplot
nl_gen <- boxplot(Total~Generation,data=NonLegendary, main="Non-Legendary Dragon Boxplot for Each Generations",xlab="Generation", ylab="Total Stats")
nl_gen

#all dragon generation boxplot
all_gen <- boxplot(Total~Generation,data=DragonStats, main="Dragon Boxplot for Each Generations",xlab="Generation", ylab="Total Stats")
all_gen
```
What I found interesting here is that there were no legendary dragon pokemons before the 3rd generation, this cause me to be curious about whether or not the generations have a big impact on the total stats. In order to test this I will create a scatter plot and create a regression line on it to see if there is any relationship between generation and total stats.

```{r relationgenstats}
nonvslegend <- DragonStats %>%
  ggplot(mapping = aes(x=Generation,y=Total)) +
  geom_point(aes(color = Legendary)) +
  geom_smooth(method = lm) +
  ggtitle("Total Stats Across Generations") + 
  ylab("Total Stats") + xlab("Generation")
nonvslegend


lgm <- lm(Total~Generation, data=DragonStats)
broom::tidy(lgm)
```

This regression line is suprising, it actually shows a slightly negative relationship between generations and total stats. As the generations increases the total stats of dragon type pokemon decreases. I was expecting it to increase. I will now clean up the data a bit more to see if the relationship remains the same. 
```{r cleanup}
cleaned <- DragonStats %>%
  group_by(Generation) %>%
  summarize(min_stat=min(Total), mean_stat=mean(Total), max_stat=max(Total) , sd_stat = sd(Total))
cleaned

Generation_plot <- ggplot(DragonStats, aes(x = Name, y = Total, colour = Generation)) + geom_point() +
  facet_wrap( ~ Generation)

Generation_plot
```

From the tables and the plots above it suggests that a linear regression model may not be appropriate for this data. So what I will do now is check the residual plots to see how well the linear regression model fits the data. 

```{r residualplots}
lgm %>%
  broom::augment() %>%
  ggplot(aes(x=factor(Generation), y=.resid)) +
    geom_violin() +
    labs(title="Residuals vs Generations",
         x = "Generations",
         y = "Residuals")
```
From observing the residual violins we can tell that the residuals are not zero centered and the distributions are all weird, some are unimodal some are bimodal and some are neither. 

We will now use a 5-fold cross validation to see how well the linear regression line predicts  total stats based on generations. We are using a 5-fold cross validation because our data frame is small for just dragon type pokemons. 
You can learn more about cross validation by visting this link here:
http://www.hcbravo.org/IntroDataSci/bookdown-notes/model-selection.html
and reading lecture 32.1
For my cross validation I will be utilizing the caret library as well as the psych library.
You can get more information on the caret library such as documentations by visiting this link: https://cran.r-project.org/web/packages/caret/caret.pdf 
You can get more information on the psych library by visiting this link:
https://cran.r-project.org/web/packages/psych/psych.pdf

```{r crossvalidation}
library(caret)
library(psych)
#doing a 5 fold cross validation
data_train <- trainControl(method = "cv" , number = 5)
caret_model <- train(Total ~ Generation,
                     data = DragonStats,
                     tr = data_train,
                     method = "lm",
                     na.action = na.pass) 
# this is the model prediction
# notice how the Rsquared value is very low at only 0.0519049 which means
# that only approximately 5.19% of the total stats variation is explained by Generations thus a linear model is a very poor fit for this data.
caret_model
# we can examine the predicitons for each fold of the model
caret_model$resample
```
Now we can compare the 5 cross fold r-square values against the r-square value of the whole data. The r-squared value when using the whole set of data is 0.02258 which is even lower than when using the 5 cross fold, this indicates that splitting my data into training and test sets provided a better method for getting model estimations for total stats based on generations of pokemons.
```{r summarystats}
summary_stats <- lm(Total ~ Generation, data = DragonStats)
summary(summary_stats)
```


## Insights
Now that we covered how to create graphs using a data set after tidying up the data frame, make hypothesis tests to answer questions you make be curious regarding the data, and using residual plots and 5-fold cross validation to check how well a prediction model is for the data, it is time to finally answer the question of what is the best dragon pokemon team. I will create a few data frames that will help you create the best dragon pokemon trainer based on your preferences. 
```{r bestteams}
# Dragon Pokemon Team that includes all generations as well as legendaries
StrongAll <- DragonStats %>%
  arrange(desc(Total)) %>%
  head(6)
StrongAll

# Dragon Pokemon Team that includes all generations but no legendaries
StrongNoLegend <- NonLegendary %>%
  arrange(desc(Total)) %>%
  head(6)
StrongNoLegend

# Dragon Pokemon Team that includes all generations with just legendaries
StrongLegend <- LegendaryDragons %>%
  arrange(desc(Total)) %>%
  head(6)
StrongLegend

# Dragon Pokemon Team that includes only first generation pokemons
# Note there's only 4 pokemons so you will have a disadvantage of not having a full team if you just want to have only a first generation team
StrongFirst <- DragonStats %>%
  arrange(desc(Total)) %>%
  filter(Generation == "1") %>%
  head(6)
StrongFirst

# Dragon Pokemon Team that includes only second generation pokemons
StrongSec <- DragonStats %>%
  arrange(desc(Total)) %>%
  filter(Generation == "2") %>%
  head(6)
StrongSec

# Dragon Pokemon Team that includes only third generation pokemons
StrongThird <- DragonStats %>%
  arrange(desc(Total)) %>%
  filter(Generation == "3") %>%
  head(6)
StrongThird

# Dragon Pokemon Team that includes only fourth generation pokemons
StrongFourth <- DragonStats %>%
  arrange(desc(Total)) %>%
  filter(Generation == "4") %>%
  head(6)
StrongFourth

# Dragon Pokemon Team that includes only fifth generation pokemons
StrongFifth <- DragonStats %>%
  arrange(desc(Total)) %>%
  filter(Generation == "5") %>%
  head(6)
StrongFifth

# Dragon Pokemon Team that includes only sixth generation pokemons
StrongSixth <- DragonStats %>%
  arrange(desc(Total)) %>%
  filter(Generation == "6") %>%
  head(6)
StrongSixth
```
Now we have all the teams, we can also check which generation team is the strongest. 
We can answer this question by creating a simple dataframe. 
```{r whichgeneration}
# Calculate the team stat total for each generation
sum_first <- StrongFirst %>%
  group_by(Generation) %>%
  summarize(Team_Stats = sum(Total))

sum_sec <- StrongSec %>%
  group_by(Generation) %>%
  summarize(Team_Stats = sum(Total))

sum_third <- StrongThird %>%
  group_by(Generation) %>%
  summarize(Team_Stats = sum(Total))

sum_fourth <- StrongFourth %>%
  group_by(Generation) %>%
  summarize(Team_Stats = sum(Total))

sum_fifth <- StrongFifth %>%
  group_by(Generation) %>%
  summarize(Team_Stats = sum(Total))

sum_sixth <- StrongSixth %>%
  group_by(Generation) %>%
  summarize(Team_Stats = sum(Total))

# use a full join to combine them into a single data frame
j_one <- sum_first %>%
  full_join(sum_sec, by = c( "Generation", "Team_Stats") )
j_two <- j_one %>%
  full_join(sum_third, by = c( "Generation", "Team_Stats") )
j_three <- j_two %>%
  full_join(sum_fourth, by = c( "Generation", "Team_Stats") )
j_four <- j_three %>%
  full_join(sum_fifth, by = c( "Generation", "Team_Stats") )
j_five <- j_four %>%
  full_join(sum_sixth, by = c( "Generation", "Team_Stats") )

j_five

```

So in conclusion the fourth and fifth generation has the strongest dragon type pokemon team. From creating this tutorial I was able to give a brief introduction some of the basic features of r such as Data Curation, Parsing and Management, Exploratory Data Analysis, Hypothesis Testing and Machine Learning. I also learned a lot from creating this tutorial as well such as that the fourth and fifth generation had the strongest dragon type pokemons in general. I always assumed pokemons get stronger with each new generation however my assumption was debunked when I tried to apply to a linear regression model on the data and I failed to reject the null hypothesis. I hope this tutorial was as interesting for you as it was for me! 
